### 2010-08-06 Thesis/Chapter3.tex
- Update 6 August -
git-svn-id: svn+ssh://scm.gforge.inria.fr/svnroot/sofa-shells@143 7b2761c9-2758-49ef-8c29-9c9f1e1aad1b

	\subsection{Dynamic system of equations}	
The space-time decoupled finite element formulation of time-dependent problems involves two steps:
\begin{description}
\item[1. Spatial approximation.]
The solution $ u $ is first approximated by expressions of the form \eqref{chap3:approxTime2} while keeping the time-dependent term during the finite element model derivation. This step leads to a set of ordinary differential equations in time.
\item[2. Temporal approximation.]
The system of ordinary differential equations is then approximated in time where different schemes may be used to assess the time derivatives. This step yields a set of algebraic equations for $ u^e_i $ at discrete time $ t_{k+1} = (k+1) \Delta t$ where $ \Delta t $ is the time increment and $ k > 0$ an integer. 
\end{description}
At the end of this two-step approximation, we have a continuous spatial solution at discrete intervals in time:
\begin{equation}
\label{chap3:approxTime}
u(x, t_k) \approx u^e(x, t_k) = \sum_{i=1}^n u^e_i(t_k) \psi^e_i(x) \quad \text{with } k = 0, 1, \ldots
\end{equation}

@@

	\subsection{Static system of equations}	
Under some assumption, \eqref{chap3:eqDynamic} may be simplified. For instance, let us consider the problem of finding the maximum load that a desk can sustain. It is fair to say that strain and stress within the structure does not fluctuate over time. Therefore, we can assume that the equations describing the structure do not depend on time: the problem is said to be \emph{static}. In this case, the static system of equations can be easily obtained by merely dropping out the dynamics terms in \eqref{chap3:eqDynamic}. Indeed, the static case may just be seen as a special case of the dynamic equations and the static equations are:
\begin{equation}
\label{chap3:eqStatic}
\mathbf{K}(\mathbf{U}) \cdot \mathbf{U} = \mathbf{R}.
\end{equation}

@@

		\subsubsection{Mass matrix}
The mass matrix $ \mathbf{M} $ for the whole domain may be computed by summing the mass contributions from every elements:
\begin{equation}
\mathbf{M} = \sum_e \mathbf{M}^e,
\end{equation}
where $ \mathbf{M}^e $ is the mass of element $ \Omega_e $ given by:
\begin{equation}
\mathbf{M}^e = \int_{V^e} \rho^e \mathbf{H}^T \mathbf{H} dV
\end{equation}
where $ \rho $ the mass density, $ V^e $ the volume of the element and $ \mathbf{H} $ is the shape function matrix defined by \eqref{chap3:H}. This integral may be evaluated using a procedure of numerical integration and generally results in a band mass matrix (non-zero entries are confined to a diagonal band, comprising the main diagonal and more diagonals on either side). This method for computing the mass matrix is called \emph{variational mass lumping}.

@@

	\subsection{Static problems}

	\subsection{Dynamic problems}

>>>>

	\subsection{Dynamic system of equations}	 \label{chap3:dynamic}
The space-time decoupled finite element formulation of time-dependent problems involves two steps:
\begin{description}
\item[1. Spatial approximation.]
The solution $ u $ is first approximated by expressions of the form \eqref{chap3:approxTime2} while keeping the time-dependent term during the finite element model derivation. This step leads to a set of ordinary differential equations in time.
\item[2. Temporal approximation.]
The system of ordinary differential equations is then approximated in time where different schemes may be used to assess the time derivatives. This step yields a set of algebraic equations for $ u^e_i $ at discrete time $ t_{k+1} = (k+1) \Delta t$ where $ \Delta t $ is the time increment and $ k > 0$ an integer. 
\end{description}
At the end of this two-step approximation, we have a continuous spatial solution at discrete intervals in time:
\begin{equation}
\label{chap3:approxTime}
u(x, t_k) \approx u^e(x, t_k) = \sum_{i=1}^n u^e_i(t_k) \psi^e_i(x) \quad \text{with } k = 0, 1, \ldots
\end{equation}

@@

	\subsection{Static system of equations}	\label{chap3:static}
Under some assumption, \eqref{chap3:eqDynamic} may be simplified. For instance, let us consider the problem of finding the maximum load that a desk can sustain. It is fair to say that strain and stress within the structure does not fluctuate over time: the system is in an equilibrium state. At all times, the sum of all forces applied on the system is equal to $ 0 $. Therefore, we can assume that the equations describing the structure do not depend on time: the problem is said to be \emph{static}. In this case, the static system of equations can be easily obtained by merely dropping out the dynamics terms in \eqref{chap3:eqDynamic}. Indeed, the static case may just be seen as a special case of the dynamic equations and the static equations are:
\begin{equation}
\label{chap3:eqStatic}
\mathbf{K}(\mathbf{U}) \cdot \mathbf{U} = \mathbf{R}.
\end{equation}

@@

		\subsubsection{Mass matrix} \label{chap3:massMatrix}
The mass matrix $ \mathbf{M} $ for the whole domain may be computed by summing the mass contributions from every elements:
\begin{equation}
\mathbf{M} = \sum_e \mathbf{M}^e,
\end{equation}
where $ \mathbf{M}^e $ is the mass of element $ \Omega_e $ given by:
\begin{equation}
\mathbf{M}^e = \int_{V^e} \rho^e \mathbf{H}^T \mathbf{H} dV
\end{equation}
where $ \rho $ the mass density, $ V^e $ the volume of the element and $ \mathbf{H} $ is the shape function matrix defined by \eqref{chap3:H}. This integral may be evaluated using a procedure of numerical integration and generally results in a band mass matrix (non-zero entries are confined to a diagonal band, comprising the main diagonal and more diagonals on either side). This method for computing the mass matrix is called \emph{variational mass lumping}.

@@

\cite{Belytschko00}

	\subsection{Explicit time integration}

	\subsection{Static solutions and implicit time integration}

%The complete and global system of equations has now been established. In this section, we will examine the techniques commonly used for solving these equations. Note that this is only the most classical approaches and the list is obviously not exhaustive. The choice of a resolution method depends on the problem (linear or non-linear) and the properties of the stiffness matrix. The system will first be solved in static before reviewing different methods of time integration to find the solutions in dynamic. 
%
%	\subsection{Static problems}
%First, we recall the general system of equations in static introduced in section \ref{chap3:static}, which is also valid for the whole system: 
%\begin{equation}
%\mathbf{K}(\mathbf{U}) \cdot \mathbf{U} = \mathbf{R}.
%\end{equation}	
%In this case, the displacement $ \mathbf{U} $ at each node may be found using an appropriate solver. If the problem is linear, the stiffness matrix does not depend on the displacement and the resolution is facilitated. The unknown displacements may be computed by
%\begin{equation}
%\mathbf{U} = \mathbf{K}^{-1} \, \mathbf{R}.
%\end{equation}
%Many techniques are available in the literature to invert a matrix and a few of them are introduced in section \ref{chap3:solvers}. 
%	
%	\subsection{Dynamic problems}
%Let us recall the general system of equations in dynamic introduced in section \ref{chap3:dynamic}, which is also valid for the whole system: 
%\begin{equation}
%\label{chap3:dynamic2}
%\mathbf{M} \mathbf{\ddot U} + \mathbf{D} \mathbf{ \dot U} + \mathbf{K}(\mathbf{U}) \cdot \mathbf{U} = \mathbf{R}.
%\end{equation}	
%In the general case of a dynamic system of equations, the equations are not yet algebraic and still depends on derivatives of the unknown displacements with respect to time. Therefore, an additional step needs to be performed before using a solver by integrating the equations over time. The different methods of time integration are classified into two main schemes: \emph{explicit} or \emph{implicit} integration. If positions and velocities are extrapolated from their values obtained at the previous time step, then this approach is qualified of explicit. Conversely, if positions and velocities are considered as unknowns for the current time step, the scheme is called implicit. The choice of one scheme over the other impacts the stability and the computational time. 
%
%		\subsubsection*{Explicit time integration}
%Explicit schemes make use	 of data from the previous time step to assess the ones from current time step. From \eqref{chap3:dynamic2} we have:
%\begin{align}
%\mathbf{M} \mathbf{\ddot U} &= \mathbf{R} - (\mathbf{D} \mathbf{ \dot U} + \mathbf{K}(\mathbf{U}) \cdot \mathbf{U}) \notag \\
%\mathbf{M} \mathbf{\ddot U} &= \mathbf{F}_{\textbf{ext}} - \mathbf{F}_{\textbf{int}}(\mathbf{U}, \mathbf{\dot U})
%\end{align}
%where $ \mathbf{F}_{\textbf{int}} $ and $ \mathbf{F}_{\textbf{ext}} $ are the internal and external forces, respectively. We denote $ \mathbf{U}_i - \mathbf{U}_{i-1} = \Delta U = h \mathbf{\dot U}_{i-1} $ and $ \mathbf{\dot U}_i - \mathbf{\dot U}_{i-1} = \Delta \dot U = h \mathbf{\ddot U}_{i-1} $ and we may write:
%\begin{equation}
%	\begin{pmatrix}
%		\Delta U \\
%		\Delta \dot U
%	\end{pmatrix}
%	= 
%	h 
%	\begin{pmatrix}
%		\mathbf{\dot U}_{i-1} \\
%		\mathbf{M}^{-1} \, (\mathbf{F}_{\textbf{ext}} - \mathbf{F}_{\textbf{int}}(\mathbf{U}, \mathbf{\dot U}))
%	\end{pmatrix}
%\end{equation}
%We note that this scheme does not require any resolution of system if the mass matrix was diagonalised by mass lumping as explained in section \ref{chap3:massMatrix}. It is therefore a particulary efficient algorithm. However, this kind of scheme is only conditionally stable and requires small time steps. Thus, the gain in computation must be balanced against the need to calculating more steps. This lack of stability increases with smaller elements and stiffer materials.

%		\subsubsection*{Implicit time integration}

	\subsection{Solvers}	\label{chap3:solvers}

		\subsubsection*{Direct solvers}
		\subsubsection*{Iterative solvers}
